---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(dplyr)
library(xtable)
data<-read.csv("activity.csv", header=TRUE, stringsAsFactors=FALSE)
data<-tbl_df(data)
```


## What is mean total number of steps taken per day?


1. Filter out NA values in the steps.
```{r}
cleanDF<-filter(data, !is.na(steps))
```

2. Calculate the total steps grouped by Date.
```{r}
sumdf<-cleanDF %>% group_by(date) %>% summarize(DailySteps=sum(steps))
```

3. Plot the Histogram of total steps per day.
```{r histogram}
hist(sumdf$DailySteps, xlab="Daily_Steps", main="Histogram of Daily Steps")
```

4. Calculating mean and median of total steps per day 

- For this exercise we'll need to remove values with 0 steps.

```{r, results='asis'}
mean_median_df<-cleanDF %>% group_by(date) %>% filter(steps>0) %>% summarize(mean_steps=mean(steps), median_steps=median(steps))
xt<-xtable(mean_median_df)
print(xt, type="html", include.rownames=FALSE)
```

## What is the average daily activity pattern?

1. Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
avg_daily_df<-cleanDF %>% group_by(interval) %>% summarize(mean_steps=mean(steps))
plot(avg_daily_df$interval, avg_daily_df$mean_steps, type="l",ylab="Avg Daily Steps", xlab="Interval")
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
