---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data

  The below code loads the libraries for ```dplyr``` and ```xtable```. Then loads the data file and converts it into dplyr dataframe.
```{r loadData, message=FALSE}
library(dplyr)
library(xtable)

data<-read.csv("activity.csv", header=TRUE, stringsAsFactors=FALSE)
data<-tbl_df(data)
```


## What is mean total number of steps taken per day?


__1__. Filter out NA values in the steps.
```{r cleanData}
cleanDF<-filter(data, !is.na(steps))
```

__2__. Calculate the total steps grouped by Date.
```{r DailySum}
DailySumdf<-cleanDF %>% group_by(date) %>% summarize(TotalSteps=sum(steps))
```

__3__. Plot the Histogram of total steps per day.
```{r histogram1}
hist(DailySumdf$TotalSteps, xlab="Daily_Steps", main="Histogram of Daily Steps")
```

__d__. Calculating mean and median of total steps per day 


```{r meansteps1, results='asis'}

mean_median_df<-DailySumdf %>% summarize(mean_steps=mean(TotalSteps), median_steps=median(TotalSteps))

xt<-xtable(mean_median_df)
print(xt, type="html", include.rownames=FALSE)
```


## What is the average daily activity pattern?

__1__. Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r plotDailyActivity1}
avg_daily_pat_df<-cleanDF %>% group_by(interval) %>% summarize(mean_steps=mean(steps))

plot(avg_daily_pat_df$interval, avg_daily_pat_df$mean_steps, type="l",ylab="Avg Daily Steps", xlab="Interval", main="Avg Daily Activity")
```


__2__. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

   - The interval is ```r filter(avg_daily_pat_df, mean_steps==max(mean_steps)) %>% select(interval)```.
   - This can be calculated as such
```{r maxstepsInterval}
  filter(avg_daily_pat_df, mean_steps==max(mean_steps)) %>% select(interval)
```

## Imputing missing values

__1__. To compute Total rows with NA's.

   - We use the original dataset ```data```. then use  dplyr function ```filter```.

```{r NArows}
naData<- filter(data, is.na(steps))
dim(naData)[1]
```
  Total Rows having NAs : ```r dim(naData)[1]```



__2__. In order to fill in missing (NA) values with some meaning full data, we'll use mean steps for each of the corresponding 5-minute interval.

   - The below code joins the original dataset `data` (having NA's) to mean steps per Interval (`avg_daily_pat_df`).
   - Then for the NA values, updates the "steps" variable in original dataset to the "mean_steps".
   - The final output with NA's filled-in is stored in `data2`.

```{r imputingNA}
data2<-left_join(data, avg_daily_pat_df, by=c("interval"="interval")) %>% 
          mutate(steps=ifelse(is.na(steps), mean_steps,steps)) %>% 
          select(steps, date, interval)
```


__3__. Lets calculate the mean and median of Total Steps per day for the new dataset.

```{r meanSteps2, results='asis'}
DailySumdf2<-data2 %>% group_by(date) %>% summarize(TotalSteps=sum(steps))

mean_median_df2<-DailySumdf2 %>% summarize(mean_steps=mean(TotalSteps), median_steps=median(TotalSteps))

xt<-xtable(mean_median_df2)
print(xt, type="html", include.rownames=FALSE)
```


__4__. Histogram


```{r histogram2}

hist(DailySumdf2$TotalSteps, xlab="Daily_Steps", main="Histogram of Daily Steps (Imputing NA")
```

The mean and median from previous analysis isnt much different from the mean and median after imputing the NA values.



## Are there differences in activity patterns between weekdays and weekends?

  - To generate the plot we add another variable to the ```data2``` dataset ```wkday``` to indicate weekday or weekend.


```{r avgDailyActivity2}
library(lubridate)
data2<-mutate(data2, wkday=ifelse(wday(ymd(date))==7 | wday(ymd(date))==1, "weekend","weekday"))

avg_daily_pat_df2<-data2 %>% 
                  group_by(interval, wkday) %>% 
                  summarize(mean_steps=mean(steps))
```

  - Based on the below plot we can conclude that there is significant difference in activity levels between weekday and weekend.

```{r plotDailyActivity2}
library(ggplot2)
g<-ggplot(avg_daily_pat_df2, aes(interval,mean_steps)) + 
    geom_line(colour="blue",size=1) + 
    facet_wrap(~wkday, ncol=1) + 
    labs(title="Avg Daily Activity (Weekday v/s Weekend)", x="Interval", y="Avg Number of Steps")
print(g)

```
