---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data

  The below code loads the libraries for ```dplyr``` and ```xtable```. Then loads the data file and converts it into dplyr dataframe.
```{r}
library(dplyr)
library(xtable)
library(lubridate)
library(ggplot2)
data<-read.csv("activity.csv", header=TRUE, stringsAsFactors=FALSE)
data<-tbl_df(data)
```


## What is mean total number of steps taken per day?


a. Filter out NA values in the steps.
```{r}
cleanDF<-filter(data, !is.na(steps))
```

b. Calculate the total steps grouped by Date.
```{r}
DailySumdf<-cleanDF %>% group_by(date) %>% summarize(TotalSteps=sum(steps))
```

c. Plot the Histogram of total steps per day.
```{r histogram}
hist(DailySumdf$TotalSteps, xlab="Daily_Steps", main="Histogram of Daily Steps")
```

d. Calculating mean and median of total steps per day 


```{r, results='asis'}

## mean_median_df<-cleanDF %>% group_by(date) %>% filter(steps>0) %>% summarize(mean_steps=mean(steps), median_steps=median(steps))

mean_median_df<-DailySumdf %>% summarize(mean_steps=mean(TotalSteps), median_steps=median(TotalSteps))

xt<-xtable(mean_median_df)
print(xt, type="html", include.rownames=FALSE)
```

## What is the average daily activity pattern?

a. Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
avg_daily_pat_df<-cleanDF %>% group_by(interval) %>% summarize(mean_steps=mean(steps))
plot(avg_daily_pat_df$interval, avg_daily_pat_df$mean_steps, type="l",ylab="Avg Daily Steps", xlab="Interval")
```

## Imputing missing values

a. To compute Total rows with NA's 

```{r}
naData<- filter(data, is.na(steps))
dim(naData)[1]
```
  Total Rows having NAs : 2304



b. In order to fill in missing (NA) values with some meaning full data, we'll use mean steps for each of the corresponding 5-minute interval.
   - The below code joins the original dataset (having NA's) to mean steps per Interval (```avg_daily_pat_df```).
   - Then for the NA values, updates the "steps" variable in original dataset to the "mean_steps".
   - The final output with NA's filled-in is stored in data2.

```{r}
data2<-left_join(data, avg_daily_pat_df, by=c("interval"="interval")) %>% mutate(steps=ifelse(is.na(steps), mean_steps,steps)) %>% select(steps, date, interval)
```


c. Lets calculate the mean and median of Total Steps per day for the new dataset.

```{r results='asis'}
DailySumdf2<-data2 %>% group_by(date) %>% summarize(TotalSteps=sum(steps))
mean_median_df2<-DailySumdf2 %>% summarize(mean_steps=mean(TotalSteps), median_steps=median(TotalSteps))

xt<-xtable(mean_median_df2)
print(xt, type="html", include.rownames=FALSE)
```


c. Histogram


```{r}

hist(DailySumdf2$TotalSteps, xlab="Daily_Steps", main="Histogram of Daily Steps (Imputing NA")
```

The mean and median from previous analysis isnt much different from the mean and median after imputing the NA values.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
data2<-mutate(data2, wkday=ifelse(wday(ymd(date))==7 | wday(ymd(date))==1, "weekend","weekday"))
avg_daily_pat_df2<-data2 %>% group_by(interval, wkday) %>% summarize(mean_steps=mean(steps))
g<-ggplot(avg_daily_pat_df2, aes(interval,mean_steps)) + geom_line(colour="blue",size=1) + facet_wrap(~wkday, ncol=1) + labs(title="Avg Daily Activity (Weekday v/s Weekend)", x="Interval", y="Avg Number of Steps")
print(g)

```
